---
title: "COMP2501 Assignment1"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Requirements

**Submission deadline: Feb 28th, 2023 at 23:59.**

**Full mark of assignment 1: 33.**

For the following questions, please:

1. Replace all [Input here] places with your information or your answer (for multiple choice).
2. Complete the code block by adding your own code to fulfill the requirements in each question. Please use the existing code block and do not add your own code block.

Please make sure your Rmd file is a valid Markdown document and can be successfully knitted.

For assignment submission, please knit your final Rmd file into a Word document, and submit both your **Rmd** file and the knitted **Microsoft Word** document file to Moodle. You get 0 score if 1) the Rmd file you submitted cannot be knitted, and 2) you have not submitted a Word document. For each visualization question, please make sure that the generated plot is shown in-place with the question and after the code block. 

---


## Name and UID

Name: Frankline Misango Oyolo

UID: 3036050867

---


### Environmental setup 
You need to have both the `dslabs` and `dplyr` packages installed. If not yet, please run `install.packages(c("dslabs", "dplyr"))` in your R environment. If you have installed the `tidyverse` package, `dplyr` is installed by default.

```{r}
# Load the packages and dataset.
library(dslabs)
library(ggrepel)
library(ggplot2)
library(dplyr)
data("murders")
```


### Example question. Print the first 2 records of the murders dataset.

```{r}
head(murders, 2)
```


### 1. (1 points) Given matrix `mat`, print the entries of row 2, and columns 2 to 4.

```{r}

mat <- matrix(1:12,3, 4)
print(mat[2, 2:4])
```


### 2. (2 points) Write a function `compute_s_n`  that for any given n, computes the $S_n = n*\sqrt{(n+9)}*log_{10}(n)$. Print the $S_n$ with $n = 500$.

```{r}
compute_s_n <- function(n){
    print(n * sqrt(n + 9) * log(n,10))
}
compute_s_n(500)
```   


### 3. (2 points) Compute the murder rate per 100,000 people for each state and store it in an object called murder_rate. Then use logical operators to find which state has a murder rate per 100,000 people higher than 5. Find these states, print their names and murder rate per 100,000 people.

```{r}
murder_rate <- (murders$total / murders$population * 100000)
ind <- murder_rate > 5
ind_states <- murders$state[ind]
ind_rates <- murder_rate[ind]
ind_states
ind_rates





```


### 4. (2 points) For all states having a murder rate per 100,000 people higher than 5, use the barplot function to create a barplot with the x-axis being the state name, and the y-axis being the murder rate per 100,000 people of each state. (Hint: check some barplot examples at https://r-graph-gallery.com/210-custom-barplot-layout.html)

```{r}
barplot(murder_rate[ind], 
        names.arg = ind_states, 
        xlab = "States", 
        ylab = "Murder rate per 100,000 people"
        )

```


### 5. (1 points)  Examine the built-in dataset `Orange`. Which of the following is true?
- a.	Orange is tidy data: it has one observation for each row.
- b.	Orange is not tidy: we need at least one column with a character vector.
- c.	Orange is not tidy: it is a matrix instead of a data frame.
- d.	Orange is tidy data: all small datasets are tidy by definition.

Your answer is: a


### 6. (3 points) Base on the `murders` dataset, create a table called `my_states` that contains rows for states satisfying two conditions: 1) it is in either West or South, and 2) the murder rate per 100,000 people is less than 2.0. Use `select` to show only the state name, the region and the murder rate, and use top_n function to find the 3 safest states among them.

```{r}
#You need a murder rate column to select the rate hence mutate
murders <- mutate(murders, murder_rate = (murders$total / murders$population * 100000))


my_states <- filter(murders, region %in% c("West", "South") & murder_rate < 2.0)
my_states <- select(my_states, state, region, murder_rate)


top_n(my_states,3, wt=murder_rate)



```


### 7. (2 points) By using the murders dataset, compute the average murder rate per 100,000 people in the four regions respectively of the U.S., and sort the results by murder rate in ascending order.

```{r}
average_murder_rate_Region_Wise <- murders |> group_by(region) |> summarize(average_murder_rate = mean(murder_rate)) |> arrange(average_murder_rate)

print(average_murder_rate_Region_Wise)

```


### 8. (3 points) Use the ggplot2 package to create a scatterplot from the murders dataset, where the x-axis is the number of population, the y-axis is the total number of murders, and each point in the scatterplot is labeled with the state name. Please add an appropriate title, and axis labels to the plot.

```{r}
options(ggrepel.max.overlaps = Inf)

murder_plot_a <- ggplot(murders, aes(x = population, y = total, label = state)) +
  geom_point() +
  geom_text_repel ()+  
  labs(title = "Number of murders per each US state", x = "Total Population per state", y = "Total Murders per state") 

print(murder_plot_a)
```


### 9. (17 points)  Explore the tidyverse with the COVID-19 dataset (http://www.bio8.cs.hku.hk/comp2501/covid.csv), and answer the following questions.

```{r}
# Load the required packages
library(dplyr)
library("ggplot2")
```


#### a. (2 points) Read the CSV formatted dataset. Find out how many observations (rows) and variables (columns) are in the dataset. Print the names of all variables.

```{r}
covid_19_url <- "http://www.bio8.cs.hku.hk/comp2501/covid.csv"
covid_19_data <- read.csv(covid_19_url)
dim(covid_19_data)
names(covid_19_data)

```


#### b. (1 points) List the observation with the largest `Cumulative_number_for_14_days_of_COVID.19_cases_per_100000`.

```{r}
covid_19_data |> arrange(desc(Cumulative_number_for_14_days_of_COVID.19_cases_per_100000)) |>  head(1)
```


#### c. (2 points) How many unique `countriesAndTerritories` are in the dataset? How many unique `continentExp` are in the dataset?

```{r}
n_distinct(covid_19_data$countriesAndTerritories)
n_distinct(covid_19_data$continentExp)

```


#### d. (3 points) For 1) the whole dataset, 2) different `countriesAndTerritories`, and 3) different `continentExp`, compute both i) the sum of cases, and ii) the sum of deaths. Sort the results by the sum of cases descendingly. Use head() if there are too many rows in the results. 

```{r}
covid_19_data |> summarise(sum_of_cases = sum(cases), sum_of_deaths = sum(deaths)) |> arrange(desc(sum_of_cases)) |> head()

covid_19_data |> group_by(countriesAndTerritories) |> summarise(sum_of_cases = sum(cases),sum_of_deaths = sum(deaths)) |> arrange(desc(sum_of_cases)) |> head()

covid_19_data |> group_by(continentExp) |> summarise(sum_of_cases = sum(cases), sum_of_deaths = sum(deaths)) |>  arrange(desc(sum_of_cases)) |> head()


```


#### e. (2 points) Add a new column `date` with the standard date format "YYYY-MM-DD" to the data table according to the `dateRep` column. Be reminded the format of `dateRep` is "DD/MM/YYYY". Please use head() to show the result.

```{r}
covid_19_data <- covid_19_data |> mutate(date = lubridate::dmy(dateRep)) |> mutate(date = format(date, "%Y-%m-%d"))
  
head(covid_19_data)

```


#### f. (1 points) Create a scatterplot showing `cases` vs. `deaths`. Set an appropriate plot title and axis titles.

```{r}
covid_19_plot <- ggplot(covid_19_data, aes(x = cases, y = deaths)) + geom_point() + labs(title = "COVID-19 Cases vs Deaths", x = "Cases", y = "Deaths")
print(covid_19_plot)
```


#### g. (2 points) Create a line plot using data with `countriesAndTerritories=="China"`, showing date on the x-axis and cases per day on the y-axis. Set an appropriate plot title and axis titles.

```{r}
covid_19_plot_b <- covid_19_data |> filter(countriesAndTerritories == "China") |>  ggplot(aes(x = as.Date(date), y = cases)) +
  geom_line() +
  labs(title = "COVID-19 Cases per Day in China", x = "Date", y = "Cases")
print(covid_19_plot_b)
```


#### h. (2 points) Similar to above, create a line plot using the data of six countries including "China", "United_States_of_America", "United_Kingdom", "France", "Germany", and "Italy". Use different line colors for each country. Set an appropriate plot title and axis titles.

```{r}
six_countries <- c("China", "United_States_of_America", "United_Kingdom", "France", "Germany", "Italy")
six_data <- covid_19_data |>  filter(countriesAndTerritories %in% six_countries)

covid_19_plot_c <- ggplot(six_data, aes(x = as.Date(date), y = cases, color = countriesAndTerritories)) + 
  geom_line() + 
  labs(title = "Cases per Day in Six specific Countries ", x = "Date", y = "Cases")

print(covid_19_plot_c)

```


#### i. (2 points) Similar to question h, create a boxplot instead.

```{r}
covid_19_plot_d <- ggplot(six_data, aes(x = countriesAndTerritories, y = cases, fill = countriesAndTerritories)) + 
  geom_boxplot() + 
  labs(title = "Cases in Six Countries", x = "Country", y = "Cases")
print(covid_19_plot_d)
```